### 2020-06-10 writeup/proof_examples.tex
revise

\begin{proof}%[Proof of Proposition \ref{prop_var_transition}]
\todo{check!}In the setting of Proposition \ref{prop_var_transition}, the test loss is given by \eqref{eq_te_mtl_2task}. 
%the validation loss and the test error become
%\begin{align*}
%		\val(\hat{B}; w_1, w_2)
%	&=  n_1 \cdot \bignorm{\Sigma_1^{1/2}\left(\frac{w_1^2}{w_2^2} X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1}X_2^{\top}X_2\left (\beta_s - \frac{w_1}{w_2}\beta_t\right)}^2 \nonumber \\
%		&+ n_1 \sigma^2 \cdot \frac{w_1^2}{w_2^2} \bigtr{\left(\frac{w_1^2}{w_2^2}  X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-2} \left(\sigma_1^2 \frac{w_1^2}{w_2^2}  X_1^{\top}X_1 + \sigma_2^2  X_2^{\top}X_2\right)} \nonumber \\
%		&+ n_2 \cdot \frac{w_1^2}{w_2^2}\bignorm{\Sigma_2^{1/2}\left(\frac{w_1^2}{w_2^2} X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1}X_1^{\top}X_1\left(\beta_s - \frac{w_1}{w_2}\beta_t\right)}^2 \nonumber \\
%		&+ n_2 \sigma^2 \cdot \bigtr{\left(\frac{w_1^2}{w_2^2} X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-2} \left(\sigma_1^2 \frac{w_1^2}{w_2^2}  X_1^{\top}X_1 + \sigma_2^2  X_2^{\top}X_2\right)}, 
%\end{align*}
%and 
%where $\hat v=\hat{w_1}/\hat{w_2}$ is the global minimizer of $\val(\hat{B}; w_1, w_2)$. 
In the isotropic model, using again the concentration of random vector with i.i.d. entries, Lemma \ref{largedeviation}, we can rewrite $\te(\hat{\beta}_2^{\MTL})$ as
\begin{align*}
	\te(\hat{\beta}_2^{\MTL}) &=~ \hat{v}^2 \left[d^2 +\left( \hat v-1\right)^2\kappa^2\right]\bigtr{ (\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2} \cdot \left(1+\OO(p^{-1/2+\e})\right)\nonumber \\
	& + \sigma_2^2 \cdot \bigtr{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} } + (\sigma_1^2 -\sigma_2^2)  \cdot \bigtr{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-2} \hat v^2X_1^\top X_1}
\end{align*}
with high probability for any constant $\e>0$.

In the current setting, we can also show that  \eqref{hatv_add1}  holds for $\hat v$. 
Since the proof is almost the same as the one for Lemma \ref{lem_hat_v}, we omit the details. 
%$$|\hat w-1|=\OO(p^{-1}).$$
%as in \eqref{hatw_add1}. 
%We omit the details of the proof, since it is almost the same as the one in the proof of Lemma \ref{lem_hat_v}. 
Thus under the choice parameters in \eqref{choiceofpara}, $\te(\hat{\beta}_2^{\MTL}) $ can be simplified as in \eqref{simple1}: 
\be \label{simple2}
\begin{split}
\te(\hat{\beta}_2^{\MTL})&=(1+\OO(n^{-\e}))\cdot d^2 \tr\left[(X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right] \\ 
&+(1+\OO(n^{-\e}))\cdot \sigma_2^2  \bigtr{(X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} }  \\
&+(1+\OO(n^{-\e}))\cdot (\sigma_1^2-\sigma_2^2)  \bigtr{(X_1^{\top}X_1  + X_2^{\top}X_2)^{-2} X_1^\top X_1} .
\end{split}
\ee
Then we write 
$$ \te(\hat{\beta}_2^{\STL})-\te(\hat{\beta}_2^{\MTL}) =\delta_{\vari} - \delta_{\bias} - \delta_{\vari}^{(2)},$$
where 
$$\delta_{\vari}:=\sigma_2^2  \bigtr{(X_2^{\top}X_2)^{-1} }  - (1+\OO(n^{-\e}))\cdot \sigma_2^2  \bigtr{(X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} }$$
satisfies \eqref{Deltavar} but with $\sigma^2$ replaced with $\sigma_2^2$, 
$$\delta_{\bias}:=(1+\OO(n^{-\e}))\cdot d^2 \tr\left[(X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right]$$
satisfies \eqref{Deltabeta}, and 
\begin{align*}	
	\delta_{\vari}^{(2)}:=(1+\OO(n^{-\e}))\cdot (\sigma_1^2 -\sigma_2^2) \bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-2} X_1^\top X_1} .
\end{align*}
%$$	\te(\hat{\beta}_t^{\MTL}) = \left[\Delta_{\bias} + \Delta_{diff}+  \sigma_2^2 \cdot \tr(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} \right]\cdot \left(1+\OO(p^{-1/2})\right), $$
%where $\Delta_{\bias}:= \sigma_2^2 \cdot \bigtr{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} }$ has 
To estimate this new term $\delta_{\vari}^{(2)}$, we use the same arguments as in the proof of Lemma \ref{prop_model_shift_tight}: we first replace $X_1^\top X_1$ with $n_1\id$ up to some error using \eqref{eq_isometric}, and then apply Lemma \ref{lem_cov_derivative} to calcualte $\bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-2}}$. This process leads to the following estimates on $\delta_{\vari}^{(2)}$:
\be\label{Deltavar2} 
\al_-(\rho_1) - \oo(1)  \le \frac{\delta_{\vari}^{(2)}}{ \Delta_{\vari}^{(2)}} \le \al_+(\rho_1) +  \oo(1) , \ee
where 
$$ \Delta_{\vari}^{(2)}:=(\sigma_1^2 -\sigma_2^2) \frac{\rho_1 (\rho_1+\rho_2)}{(\rho_1+\rho_2-1)^3}.$$
Next we compare $\delta_{\vari}$ with $\delta_{\bias} + \delta_{\vari}^{(2)}$. Our main focus is to see how the extra $\delta_{\vari}^{(2)}$ affects the information transfer in this case.

@@

Note that the condition $\ex{\norm{\beta_1 - \beta_2}^2}< \frac{1}{2} {\sigma_2^2}  \cdot \Phi(\rho_1, \rho_2)$ for $\rho_1 > 40$ means the we have $\delta_{\vari}>\delta_{\bias} $ by Proposition \ref{prop_dist_transition}. Hence if $\sigma_1^2\le \sigma_2^2$, then $\delta_{\vari}^{(2)}<0$ and we always have $\delta_{\vari} > \delta_{\bias}+ \delta_{\vari}^{(2)}$, which gives $\te(\hat{\beta}_2^{\MTL})<\te(\hat{\beta}_2^{\STL})$.  It remains to consider the case $\sigma_1^2 \ge \sigma_2^2$.

\noindent{\bf Positive transfer.} By \eqref{Deltavar}, \eqref{Deltabeta} and \eqref{Deltavar2}, if the following inequality holds,
\be\label{cond sigma1}
\begin{split} 
&\sigma_2^2  \cdot \frac{\rho_1}{(\rho_2-1)(\rho_1 + \rho_2 -1)} \cdot (1-\oo(1)) \\
&>pd^2 \cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3}\al_+^2(\rho_1)  + (\sigma_1^2 -\sigma_2^2)\cdot \frac{\rho_1 (\rho_1+\rho_2)}{(\rho_1+\rho_2-1)^3} \al_+(\rho_1)  , 
\end{split}
\ee
then we have $\delta_{\vari} > \delta_{\bias} + \delta_{\vari}^{(2)}$ w.h.p., which gives $\te(\hat{\beta}_t^{\MTL})<\te(\hat{\beta}_t^{\STL})$. We can solve \eqref{cond sigma1} to get
\begin{align*}
\sigma_1^2 < - pd^2\cdot \rho_1 \al_+(\rho_1) +\sigma_2^2 \left[ 1+ \rho_1\Phi(\rho_1, \rho_2) \al_+^{-1}(\rho_1)\right]\cdot (1-\oo(1)).
\end{align*}
This proves the first claim of Proposition \ref{prop_var_transition} for positive transfer. \cor $\nu = (1+\oo(1)) \cdot (1 - 1/\sqrt{\rho_1})^{-4}$\nc

\noindent{\bf Negative transfer.} On the other hand, if the following inequality holds,
\be\label{cond sigma12}
\begin{split} 
&\sigma_2^2 \cdot \frac{\rho_1}{(\rho_2-1)(\rho_1 + \rho_2 -1)}\cdot \left(1 + \oo(1)\right) \\
&< pd^2 \cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3}\al_-^2(\rho_1)  + (\sigma_1^2 -\sigma_2^2)\cdot \frac{\rho_1 (\rho_1+\rho_2)}{(\rho_1+\rho_2-1)^3} \al_-(\rho_1) , 
\end{split}
\ee
then we have $\delta_{\vari} < \delta_{\bias} + \delta_{\vari}^{(2)}$ w.h.p., which gives $\te(\hat{\beta}_t^{\MTL})>\te(\hat{\beta}_t^{\STL})$. We can solve \eqref{cond sigma12} to get
\begin{align*}
\sigma_1^2 > - pd^2 \cdot \rho_1 \al_-(\rho_1) +\sigma_2^2 \left[ 1+ \rho_1\Phi(\rho_1, \rho_2)\al_-^{-1}(\rho_1)\right]\cdot (1+\oo(1)).
\end{align*}
This proves the second claim of Proposition \ref{prop_var_transition} for negative transfer.
\end{proof}

@@

\begin{proof}[Proof of Proposition \ref{prop_data_size}]
Following the above proof of Proposition \ref{prop_dist_transition}, we see that $\te(\hat{\beta}_2^{\MTL})< \te(\hat{\beta}_2^{\STL})$ w.h.p. if \eqref{pos1} holds, while $\te(\hat{\beta}_2^{\MTL})>\te(\hat{\beta}_2^{\STL})$ w.h.p. if \eqref{neg1} holds. 
%\be\label{pos1}
%pd^2  \cdot \frac{\rho_1 (\rho_1 + \rho_2) (\rho_2 - 1)}{(\rho_1 + \rho_2 - 1)^2}\bigbrace{1 + \sqrt{\frac{1}{\rho_1}}}^4 <  \sigma^2  \bigbrace{1 - \oo(1)},\ee
%and we get from \eqref{upper102} that $\te(\hat{\beta}_t^{\MTL})>\te(\hat{\beta}_t^{\STL})$ w.h.p. if
%\be\label{neg1}pd^2  \cdot \frac{\rho_1 (\rho_1 + \rho_2) (\rho_2 - 1)}{(\rho_1 + \rho_2 - 1)^2}\bigbrace{1 - \sqrt{\frac{1}{\rho_1}}}^4 >  \sigma^2  \bigbrace{1 + \oo(1)}.\ee
%then we have $\te(\hat{\beta}_t^{\MTL})> \te(\hat{\beta}_t^{\STL})$ w.h.p..

@@

\noindent{\bf Positive transfer.} With \eqref{trivialphi}, we see that \eqref{pos1} is implied by the following inequality:
\begin{align}\label{pos1solv}
 &  \Psi(\beta_1,\beta_2) \cdot \frac{(\rho_2-1) \rho_1 }{\rho_1+\rho_2-2}   < \gamma_+^{-1} .
 \end{align}
 then we can solve \eqref{pos1solv} to get
 \be\label{addconstraint}\rho_1 < \frac{\rho_2-2}{\Psi(\beta_1,\beta_2) \cdot \gamma_+ (\rho_2 - 1) - 1}  .\ee
 %$$\rho_1 > \frac{(\rho_2-2) \sigma^2}{(1 - {\rho_1}^{-0.5})^4 (\rho_2 - 3) pd^2 - \sigma^2}$$
 This gives the first statement of Proposition \ref{prop_data_size}. %Plugging into $\rho_1>40$ and $\rho_2>500$, we conclude the upper bound for $\rho_1$ in order for positive transfer to hold.

 Note that if we require the RHS of \eqref{addconstraint} to be larger than $40$, that is, \eqref{addconstraint} is not a null condition. Then together with \eqref{explainpsi}, we get
 $$ \rho_2 - 2>( 2\gamma_+ -1)\rho_1 .$$
Plugging into $\rho_1>40$, we get $\rho_2 \ge 106$. This gives a constraint on $\rho_2$.

@@

 \noindent{\bf Negative transfer.} 
With \eqref{trivialphi}, we see that \eqref{neg1} is implied by the following inequality:
\begin{align}\label{neg1solv}
 & \Psi(\beta_1,\beta_2)  \cdot \frac{(\rho_2-1) \rho_1 }{\rho_1+\rho_2-2}\left( 1-\frac{1}{(\rho_1+\rho_2-2)^2}\right) >  \Psi(\beta_1,\beta_2)  \cdot \frac{(\rho_2-1.5) \rho_1 }{\rho_1+\rho_2-2}> \gamma_-^{-1} .
 \end{align}
 where we used $(1-(\rho_1+\rho_2-2)^{-2})(\rho_2 - 1) > \rho_2-1.5$ for $\rho_1>40$ and $\rho_2>110$. Then we can solve \eqref{neg1solv} to get
 \be\label{addconstaint2}
 \rho_1 > \frac{(\rho_2-2) \sigma^2}{\Psi(\beta_1, \beta_2) \cdot \gamma_-(\rho_2 - 1.5)  - 1}  ,
 \ee
which gives the second statement of Proposition \ref{prop_data_size}.
We remark that condition \eqref{explainpsi} implies $\Psi(\beta_1, \beta_2) \cdot \gamma_-(\rho_2 - 1.5)>1$, so \eqref{addconstaint2} does not give a trivial bound. 
 \end{proof}

Next we state Proposition \ref{prop_data_efficiency}, which gives precise upper and lower bounds on the data efficiency ratio for taskonomy. %In the statement, we shall denote the data efficiency ratio as $\al^\star$.

\begin{proposition}[Labeled data efficiency]\label{prop_data_efficiency}
	In the isotropic model, assume that $\rho_1,\rho_2 \ge 9$ and $\Psi(\beta_1, \beta_2) < (5(\rho_1-1))^{-1} + (5(\rho_2-1))^{-1}$.
	Then the data efficiency ratio $x^\star$ satisfies 
	\be\label{eq_uplowx} x_l  \le x^\star\le \frac{1}{\rho_1 + \rho_2} \bigbrace{  \frac{2}{(\rho_1-1)^{-1} + (\rho_2-1)^{-1} - 5\Psi(\beta_1, \beta_2)}+1}, \ee
	where we denoted
	$$x_l:= \frac1{\rho_1+\rho_2}\left(\frac{2}{(\rho_1-1)^{-1}+(\rho_2 -1)^{-1}}+1\right).$$
\end{proposition}

>>>>

\begin{proof}%[Proof of Proposition \ref{prop_var_transition}]
In the setting of Proposition \ref{prop_var_transition}, the test loss is given by \eqref{eq_te_mtl_2task}. 
%the validation loss and the test error become
%\begin{align*}
%		\val(\hat{B}; w_1, w_2)
%	&=  n_1 \cdot \bignorm{\Sigma_1^{1/2}\left(\frac{w_1^2}{w_2^2} X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1}X_2^{\top}X_2\left (\beta_s - \frac{w_1}{w_2}\beta_t\right)}^2 \nonumber \\
%		&+ n_1 \sigma^2 \cdot \frac{w_1^2}{w_2^2} \bigtr{\left(\frac{w_1^2}{w_2^2}  X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-2} \left(\sigma_1^2 \frac{w_1^2}{w_2^2}  X_1^{\top}X_1 + \sigma_2^2  X_2^{\top}X_2\right)} \nonumber \\
%		&+ n_2 \cdot \frac{w_1^2}{w_2^2}\bignorm{\Sigma_2^{1/2}\left(\frac{w_1^2}{w_2^2} X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1}X_1^{\top}X_1\left(\beta_s - \frac{w_1}{w_2}\beta_t\right)}^2 \nonumber \\
%		&+ n_2 \sigma^2 \cdot \bigtr{\left(\frac{w_1^2}{w_2^2} X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-2} \left(\sigma_1^2 \frac{w_1^2}{w_2^2}  X_1^{\top}X_1 + \sigma_2^2  X_2^{\top}X_2\right)}, 
%\end{align*}
%and 
%where $\hat v=\hat{w_1}/\hat{w_2}$ is the global minimizer of $\val(\hat{B}; w_1, w_2)$. 
In the isotropic model, using again the concentration result, Lemma \ref{largedeviation}, we can rewrite $\te(\hat{\beta}_2^{\MTL})$ as
\begin{align*}
	\te(\hat{\beta}_2^{\MTL}) &=~ \hat{v}^2 \left[d^2 +\left( \hat v-1\right)^2\kappa^2\right]\bigtr{ (\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2} \cdot \left(1+\OO(p^{-1/2+\e})\right)\nonumber \\
	& + \sigma_2^2 \cdot \bigtr{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} } + (\sigma_1^2 -\sigma_2^2)  \cdot \bigtr{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-2} \hat v^2X_1^\top X_1}
\end{align*}
with high probability for any constant $\e>0$.

In the current setting, we can also show that  \eqref{hatv_add1}  holds for $\hat v$. 
Since the proof is almost the same as the one for Lemma \ref{lem_hat_v}, we omit the details. 
%$$|\hat w-1|=\OO(p^{-1}).$$
%as in \eqref{hatw_add1}. 
%We omit the details of the proof, since it is almost the same as the one in the proof of Lemma \ref{lem_hat_v}. 
Thus under the choice of parameters in \eqref{choiceofpara}, $\te(\hat{\beta}_2^{\MTL}) $ can be simplified as in \eqref{simple1}: 
\be \label{simple2}
\begin{split}
\te(\hat{\beta}_2^{\MTL})&=(1+\OO(n^{-\e}))\cdot d^2 \tr\left[(X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right] \\ 
&+(1+\OO(n^{-\e}))\cdot \sigma_2^2  \bigtr{(X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} }  \\
&+(1+\OO(n^{-\e}))\cdot (\sigma_1^2-\sigma_2^2)  \bigtr{(X_1^{\top}X_1  + X_2^{\top}X_2)^{-2} X_1^\top X_1} .
\end{split}
\ee
Then we have 
$$ \te(\hat{\beta}_2^{\STL})-\te(\hat{\beta}_2^{\MTL}) =\delta_{\vari} - \delta_{\bias} - \delta_{\vari}^{(2)},$$
where 
$$\delta_{\vari}:=\sigma_2^2  \bigtr{(X_2^{\top}X_2)^{-1} }  - (1+\OO(n^{-\e}))\cdot \sigma_2^2  \bigtr{(X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} }$$
satisfies \eqref{Deltavar} with $\sigma^2$ replaced by $\sigma_2^2$, 
$$\delta_{\bias}:=(1+\OO(n^{-\e}))\cdot d^2 \tr\left[(X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right]$$
satisfies \eqref{Deltabeta}, and $\delta_{\vari}^{(2)}$ is defined as
\begin{align*}	
	\delta_{\vari}^{(2)}:=(1+\OO(n^{-\e}))\cdot (\sigma_1^2 -\sigma_2^2) \bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-2} X_1^\top X_1} .
\end{align*}
%$$	\te(\hat{\beta}_t^{\MTL}) = \left[\Delta_{\bias} + \Delta_{diff}+  \sigma_2^2 \cdot \tr(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} \right]\cdot \left(1+\OO(p^{-1/2})\right), $$
%where $\Delta_{\bias}:= \sigma_2^2 \cdot \bigtr{(\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} }$ has 
To estimate this new term $\delta_{\vari}^{(2)}$, we use the same arguments as in the proof of Lemma \ref{prop_model_shift_tight}: we first replace $X_1^\top X_1$ with $n_1\id$ up to some error using \eqref{eq_isometric}, and then apply Lemma \ref{lem_cov_derivative} to calcualte $\bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-2}}$. This process leads to the following estimates on $\delta_{\vari}^{(2)}$:
\be\label{Deltavar2} 
\al_-(\rho_1) - \oo(1)  \le  {\delta_{\vari}^{(2)}}/{ \Delta_{\vari}^{(2)}} \le \al_+(\rho_1) +  \oo(1) , \ee
where 
$$ \Delta_{\vari}^{(2)}:=(\sigma_1^2 -\sigma_2^2) \frac{\rho_1 (\rho_1+\rho_2)}{(\rho_1+\rho_2-1)^3}.$$
Next we compare $\delta_{\vari}$ with $\delta_{\bias} + \delta_{\vari}^{(2)}$. Our main focus is to see how the extra $\delta_{\vari}^{(2)}$ affects the information transfer in this case.

@@

Note that the condition $\ex{\norm{\beta_1 - \beta_2}^2}< \frac{1}{2} {\sigma_2^2}  \cdot \Phi(\rho_1, \rho_2)$ for $\rho_1 > 40$ gives that $\delta_{\vari}>\delta_{\bias} $ by Proposition \ref{prop_dist_transition}. Hence if $\sigma_1^2\le \sigma_2^2$, then $\delta_{\vari}^{(2)}<0$ and we always have $\delta_{\vari} > \delta_{\bias}+ \delta_{\vari}^{(2)}$, which gives $\te(\hat{\beta}_2^{\MTL})<\te(\hat{\beta}_2^{\STL})$.  It remains to consider the case $\sigma_1^2 \ge \sigma_2^2$.

\noindent{\bf Positive transfer.} By \eqref{Deltavar}, \eqref{Deltabeta} and \eqref{Deltavar2}, if the following inequality holds,
\be\label{cond sigma1}
\begin{split} 
&\sigma_2^2  \cdot \frac{\rho_1}{(\rho_2-1)(\rho_1 + \rho_2 -1)} \cdot (1-\oo(1)) \\
&>pd^2 \cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3}\al_+^2(\rho_1)  + (\sigma_1^2 -\sigma_2^2)\cdot \frac{\rho_1 (\rho_1+\rho_2)}{(\rho_1+\rho_2-1)^3} \al_+(\rho_1)  , 
\end{split}
\ee
then we have $\delta_{\vari} > \delta_{\bias} + \delta_{\vari}^{(2)}$ w.h.p., which gives $\te(\hat{\beta}_t^{\MTL})<\te(\hat{\beta}_t^{\STL})$. We can solve \eqref{cond sigma1} to get
\begin{align*}
\sigma_1^2 < - pd^2\cdot \rho_1 \al_+(\rho_1) +\sigma_2^2 \left[ 1+ \rho_1\Phi(\rho_1, \rho_2) \al_+^{-1}(\rho_1)\right]\cdot (1-\oo(1)).
\end{align*}
This proves the first claim of Proposition \ref{prop_var_transition} using $\nu\ge \al_+^2(\rho_1) +  \oo(1) $.

\noindent{\bf Negative transfer.} On the other hand, if the following inequality holds,
\be\label{cond sigma12}
\begin{split} 
&\sigma_2^2 \cdot \frac{\rho_1}{(\rho_2-1)(\rho_1 + \rho_2 -1)}\cdot \left(1 + \oo(1)\right) \\
&< pd^2 \cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3}\al_-^2(\rho_1)  + (\sigma_1^2 -\sigma_2^2)\cdot \frac{\rho_1 (\rho_1+\rho_2)}{(\rho_1+\rho_2-1)^3} \al_-(\rho_1) , 
\end{split}
\ee
then we have $\delta_{\vari} < \delta_{\bias} + \delta_{\vari}^{(2)}$ w.h.p., which gives $\te(\hat{\beta}_t^{\MTL})>\te(\hat{\beta}_t^{\STL})$. We can solve \eqref{cond sigma12} to get
\begin{align*}
\sigma_1^2 > - pd^2 \cdot \rho_1 \al_-(\rho_1) +\sigma_2^2 \left[ 1+ \rho_1\Phi(\rho_1, \rho_2)\al_-^{-1}(\rho_1)\right]\cdot (1+\oo(1)).
\end{align*}
This proves the second claim of Proposition \ref{prop_var_transition}.
\end{proof}

@@

\begin{proof}[Proof of Proposition \ref{prop_data_size}]
Following the proof of Proposition \ref{prop_dist_transition}, we see that $\te(\hat{\beta}_2^{\MTL})< \te(\hat{\beta}_2^{\STL})$ w.h.p. if \eqref{pos1} holds, while $\te(\hat{\beta}_2^{\MTL})>\te(\hat{\beta}_2^{\STL})$ w.h.p. if \eqref{neg1} holds. 
%\be\label{pos1}
%pd^2  \cdot \frac{\rho_1 (\rho_1 + \rho_2) (\rho_2 - 1)}{(\rho_1 + \rho_2 - 1)^2}\bigbrace{1 + \sqrt{\frac{1}{\rho_1}}}^4 <  \sigma^2  \bigbrace{1 - \oo(1)},\ee
%and we get from \eqref{upper102} that $\te(\hat{\beta}_t^{\MTL})>\te(\hat{\beta}_t^{\STL})$ w.h.p. if
%\be\label{neg1}pd^2  \cdot \frac{\rho_1 (\rho_1 + \rho_2) (\rho_2 - 1)}{(\rho_1 + \rho_2 - 1)^2}\bigbrace{1 - \sqrt{\frac{1}{\rho_1}}}^4 >  \sigma^2  \bigbrace{1 + \oo(1)}.\ee
%then we have $\te(\hat{\beta}_t^{\MTL})> \te(\hat{\beta}_t^{\STL})$ w.h.p..

@@

\noindent{\bf Positive transfer.} With \eqref{trivialphi}, we see that \eqref{pos1} is implied by the following inequality:
\begin{align}\label{pos1solv}
 &  \Psi(\beta_1,\beta_2) \cdot \frac{(\rho_2-1) \rho_1 }{\rho_1+\rho_2-2}   < \left( \al_+^2(\rho_1) +  \oo(1) \right)^{-1} .
 \end{align}
 Then we can solve \eqref{pos1solv} to get
 \be\label{addconstraint}\rho_1 < \frac{\rho_2-2}{\Psi(\beta_1,\beta_2) \cdot (\rho_2 - 1)\left( \al_+^2(\rho_1) +  \oo(1) \right) - 1}  .\ee
 %$$\rho_1 > \frac{(\rho_2-2) \sigma^2}{(1 - {\rho_1}^{-0.5})^4 (\rho_2 - 3) pd^2 - \sigma^2}$$
 This gives the first statement of Proposition \ref{prop_data_size} using $\nu\ge \al_+^2(\rho_1) +  \oo(1) $. %Plugging into $\rho_1>40$ and $\rho_2>500$, we conclude the upper bound for $\rho_1$ in order for positive transfer to hold.

 Note that if we require the RHS of \eqref{addconstraint} to be larger than $40$, that is, \eqref{addconstraint} is not a null condition. Then together with \eqref{explainpsi}, we get
 $$ \rho_2 - 2>\left[ 2\left( \al_+^2(\rho_1) +  \oo(1) \right)  -1\right]\rho_1 .$$
Plugging into $\rho_1>40$, we get $\rho_2 \ge 106$. This gives a constraint on $\rho_2$.

@@

 \noindent{\bf Negative transfer.} 
With \eqref{trivialphi}, we see that \eqref{neg1} is implied by the following inequality:
\begin{align}\label{neg1solv}
 & \Psi(\beta_1,\beta_2)  \frac{(\rho_2-1) \rho_1 }{\rho_1+\rho_2-2}\left( 1-\frac{1}{(\rho_1+\rho_2-2)^2}\right) >  \Psi(\beta_1,\beta_2) \frac{(\rho_2-1.5) \rho_1 }{\rho_1+\rho_2-2}> \left( \al_-^2(\rho_1) -  \oo(1) \right)^{-1} .
 \end{align}
 where we used $(1-(\rho_1+\rho_2-2)^{-2})(\rho_2 - 1) > \rho_2-1.5$ for $\rho_1>40$ and $\rho_2>110$. Then we can solve \eqref{neg1solv} to get
 \be\label{addconstaint2}
 \rho_1 > \frac{(\rho_2-2) \sigma^2}{\Psi(\beta_1, \beta_2) \cdot  (\rho_2 - 1.5)\left( \al_-^2(\rho_1) -  \oo(1) \right)  - 1}  ,
 \ee
which gives the second statement of Proposition \ref{prop_data_size}.
We remark that condition \eqref{explainpsi} implies $\Psi(\beta_1, \beta_2) \cdot  (\rho_2 - 1.5)\left( \al_-^2(\rho_1) -  \oo(1) \right)>1$, so \eqref{addconstaint2} does not give a trivial bound. 
 \end{proof}

Next we state Proposition \ref{prop_data_efficiency}, which gives precise upper and lower bounds on the data efficiency ratio for Taskonomy. %In the statement, we shall denote the data efficiency ratio as $\al^\star$.

\begin{proposition}%[Labeled data efficiency]
\label{prop_data_efficiency}
	In the isotropic model, assume that $\rho_1,\rho_2 \ge 9$ and $\Psi(\beta_1, \beta_2) < (5(\rho_1-1))^{-1} + (5(\rho_2-1))^{-1}$.
	Then the data efficiency ratio $x^\star$ satisfies 
	\be\label{eq_uplowx} x_l  \le x^\star\le \frac{1}{\rho_1 + \rho_2} \bigbrace{  \frac{2}{(\rho_1-1)^{-1} + (\rho_2-1)^{-1} - 5\Psi(\beta_1, \beta_2)}+1}, \ee
	where we denoted
	$$x_l:= \frac1{\rho_1+\rho_2}\left(\frac{2}{(\rho_1-1)^{-1}+(\rho_2 -1)^{-1}}+1\right).$$
\end{proposition}

\begin{proof}%[Proof of Proposition \ref{prop_data_efficiency}]
Suppose we have reduced number of datapoints---$x n_1$ for task 1 and $x n_2$ for task 2 with $n_1=\rho_1 p$ and $n_2=\rho_2 p$. Then all the results in the proof of Proposition \ref{prop_dist_transition} still hold, except that we need to replace $(\rho_1,\rho_2)$ with $(x\rho_1,x\rho_2)$. More precisely, we have
	\begin{align*}
		 a_1 = \frac{\rho_1(x\rho_1 + x\rho_2 - 1)}{x(\rho_1 + \rho_2)^2} ,\quad
		& a_2 = \frac{\rho_2(x\rho_1 + x\rho_2 - 1)}{x(\rho_1 + \rho_2)^2} ,  \\
		 a_3 = \frac{\rho_2}{(\rho_1 + \rho_2)(x\rho_1 + x\rho_2 - 1)}, \quad
		& a_4 = \frac{\rho_1}{(\rho_1 + \rho_2)(x\rho_1 + x\rho_2 - 1)}. 
	\end{align*}
Moreover, with high probability,
	\be\label{reduceproof1}
\begin{split}
\te_i(\hat \beta_i^{\MTL}(x)) =  \frac{\sigma^2}{x (\rho_1+\rho_2) - 1}\left(1+ \oo(1)\right)+ \delta^{(i)}_{\bias}, \quad i=1,2.
\end{split}
\ee
 Here the model shift biases $\delta^{(i)}_{\bias}$ satisfy that  
\be\nonumber %\label{Deltabeta} 
\al_-^2(\al\rho_i) - \oo(1)  \le {\delta_{\bias}^{(i)}}/{ \Delta_{\bias}^{(i)}} \le \al_+^2(\al\rho_i) +  \oo(1) , \quad i=1,2,\ee	
where $ \Delta^{(i)}_{\bias}$ are defined as 
\be \nonumber
\Delta^{(i)}_{\bias} := pd^2 \frac{(x\rho_i)^2\cdot x (\rho_1+\rho_2)}{[x ( \rho_1+\rho_2) - 1]^3} ,\quad i=1,2,. 
\ee
On the other hand, using Lemma \ref{lem_minv} we have w.h.p.,
\be\label{reduceproof2} 
\te_i(\hat{\beta}_i^{\STL}) = \frac{\sigma^2}{\rho_i -1} \left( 1+ \oo(1)\right),\quad i=1,2.
\ee
Comparing \eqref{reduceproof1} and \eqref{reduceproof2}, we immediately obtain the lower bound $x^\star\ge x_l $. 
%where
%$$\al_l:=\frac1{\rho_1+\rho_2}\left(\frac{2}{\frac1{\rho_1-1}+\frac1{\rho_2 -1}}+1\right) \ge \frac{\min(\rho_1,\rho_2)}{\rho_1+\rho_2}.$$
In fact, one can see that if $x< x_l$, then we have 
$$ \frac{2\sigma^2}{x (\rho_1+\rho_2) - 1} > \frac{\sigma^2}{\rho_1-1}+\frac{\sigma^2}{\rho_2-1},$$
that is, $\te_1(\hat{\beta}_1^{\MTL}(x)) + \te_2(\hat{\beta}_2^{\MTL}(x))$ is larger than $\te_1(\hat{\beta}_t^{\STL}) + \te_2(\hat{\beta}_t^{\STL})$ even if we do not take into account the model shift bias terms $ \delta^{(i)}_{\bias}$.

Then we try to obtain an upper bound on $x^\star$. In the following discussions, we only consider $x$ such that $x> x_l$. In particular, we have $x\rho > x_l \rho \ge \min(\rho_1,\rho_2)$, where we abbreviated $\rho:=\rho_1+\rho_2$.
